Some thoughts on Zermelo's iteration process in the context of the hyper2 package

### 2023-10-30 inst/f1points.Rmd
consistentification

(observe how this likelihood function, while informative about
$p_4+p_5$, is uninformative about $p_4\left|p_4+p_5\right.$).  We now
use a technique due to \cite{hankin2010,hankin2017} and introduce
fictional (reified) entities whose nonzero Bradley-Terry strength
helps certain competitors or sets of competitors under certain
conditions.  The canonical example would be the home-ground advantage
in association football.  If players (teams) $1,2$ with strengths
$p_1,p_2$ compete, and if our observation were $a$ home wins and $b$
away wins for team $1$, and $c$ home wins and $d$ away wins for team
$2$, then a suitable likelihood function would be

@@

where $p_H$ is a quantification of the beneficial home ground effect.
Similar techniques have been used to account for the first-move
advantage in chess, and effective coordination between members of
doubles tennis teams; we may use a similar device to account for pole
position in Formula 1.  Here I analyse seasons 2016-2019 using the
`hyper2` package \citep{hankin2017} which implements the Plackett-Luce
likelihood function with additional reified entities.

>>>>

(observe how this likelihood function, while informative about
$p_4+p_5$, is uninformative about $p_4\left|p_4+p_5\right.$).  We now
use a technique due to \cite{hankin2010,hankin2017a} and introduce
fictional (reified) entities whose nonzero Bradley-Terry strength
helps certain competitors or sets of competitors under certain
conditions.  The canonical example would be the home-ground advantage
in association football.  If players (teams) $1,2$ with strengths
$p_1,p_2$ compete, and if our observation were $a$ home wins and $b$
away wins for team $1$, and $c$ home wins and $d$ away wins for team
$2$, then a suitable likelihood function would be

@@

where $p_H$ is a quantification of the beneficial home ground effect.
Similar techniques have been used to account for the first-move
advantage in chess, and effective coordination between members of
doubles tennis teams; we may use a similar device to account for pole
position in Formula 1.  Here I analyse seasons 2016-2019 using the
`hyper2` package \citep{hankin2017a} which implements the Plackett-Luce
likelihood function with additional reified entities.

### 2023-10-30 inst/mann_whitney_wilcoxon.Rmd
consistentification

---
title: "Mann-Whitney-Wilcoxon test with `hyper3` objects"
author: "Robin KS Hankin"
date: "10/04/2022"
output: html_document
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

We do some work with Mann-Whitney test, using the chorioamnion dataset at `wilcox.test.Rd`:

@@

>>>>

---
title: "Generalizing the Mann-Whitney-Wilcoxon test with `hyper3` objects"
author: "Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I discuss the `hyper2` generalization of the
Mann-Whitney-Wilcoxon test, applying it to the chorioamnion dataset at
`wilcox.test.Rd`:

@@

# References

### 2023-10-30 inst/rowing.Rmd
consistentification

---
title: "rowing: analysis of sculls in the 2016 Olympics"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document analyses the `rowing` dataset: sculls in the 2016
olympics.  It is not really standalone and is here as part of a
consistent dataset generation mechanism used in the `hyper2` package.
See the `hyper2` vignette for a more formal discussion.  In Olympic
sculling, up to six individual competitors race a small boat called a
scull over a course of length $2\,\mathrm{km}$; the object is to cross
the finishing line first.  Note that actual timings are irrelevant,
given the model, as the sufficient statistic is the order in which
competitors cross the finishing line.  The 2016 Summer Olympics is a
case in point: the gold and silver medallists finished less than $\sim
5$ milliseconds apart, corresponding to a lead of $\sim
2.5\,\mathrm{cm}$.  The package dataset is file `rowing.txt`, here are
the first three lines:

@@

>>>>

---
title: "rowing: analysis of sculls in the 2016 Olympics"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document analyses the `rowing` dataset: sculls in the 2016
olympics.  It is not really standalone and is here as part of a
consistent dataset generation mechanism used in the `hyper2` package.
See the `hyper2` vignette for a more formal discussion.  In Olympic
sculling, up to six individual competitors race a small boat called a
scull over a course of length $2\,\mathrm{km}$; the object is to cross
the finishing line first.  Note that actual timings are irrelevant,
given the model, as the sufficient statistic is the order in which
competitors cross the finishing line.  The 2016 Summer Olympics is a
case in point: the gold and silver medallists finished less than $\sim
5$ milliseconds apart, corresponding to a lead of $\sim
2.5\,\mathrm{cm}$.  The package dataset is file `rowing.txt`, here are
the first three lines:

@@

## References {-}

### 2023-10-30 inst/salad.Rmd
consistentification

---
title: "salad"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document shows how to apply `hyper2` idiom to the `salad`
dataset of the `prefmod` package.  From `salad.Rd`:

>>>>

---
title: "salad"
output: bookdown::html_document2
author: "R. K. S. Hankin"
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document shows how to apply `hyper2` idiom to the `salad`
dataset of the `prefmod` package.  From `salad.Rd`:

### 2023-10-30 inst/skating.Rmd
consistentification

---
title: "Ladies' figure skating in the 2002 Winter Olympics"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```
<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This document analyses Bradley-Terry strengths of the competitors in
the Ladies' Figure Skating at the 2002 Winter Olympics.  In the
package, dataframe `skating_table` is copied from Lock and Lock (see
skating.Rd for more details).  It also creates file `data/skating.rda`
which contains R objects `skating`, `skating_table`, and
`skating_maxp`.

@@

>>>>

---
title: "Ladies' figure skating in the 2002 Winter Olympics"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This document analyses Bradley-Terry strengths of the competitors in
the Ladies' Figure Skating at the 2002 Winter Olympics.  In the
package, dataframe `skating_table` is copied from Lock and Lock (see
skating.Rd for more details).  It also creates file `data/skating.rda`
which contains R objects `skating`, `skating_table`, and
`skating_maxp`.

@@

### References {-}

### 2023-10-30 inst/skeleton.Rmd
consistentification

---
title: "skeleton"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

The [skeleton](https://en.wikipedia.org/wiki/Skeleton_(sport)) is a
winter sliding sport.  This short document discusses a dataset taken
from the [2018 Winter
Olympics](https://en.wikipedia.org/wiki/Skeleton_at_the_2018_Winter_Olympics_%E2%80%93_Men%27s)
and analyses the order statistic for the competitors.

>>>>

---
title: "skeleton racing with the hyper2 package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
The [skeleton](https://en.wikipedia.org/wiki/Skeleton_(sport)) is a
winter sliding sport.  This short document discusses a dataset taken
from the [2018 Winter
Olympics](https://en.wikipedia.org/wiki/Skeleton_at_the_2018_Winter_Olympics_%E2%80%93_Men%27s)
and analyses the order statistic for the competitors.

### 2023-10-30 inst/soling.Rmd
add hex sticker to more Rmd files
consistentification

---
title: "soling"
output: bookdown::html_document2
---

@@

The Soling is a type of boat raced in the 2000 Summer Olympics.  Here
I analyse race data taken from
[wikipedia](https://en.wikipedia.org/w/index.php?title=Sailing_at_the_2000_Summer_Olympics_\%E2\%80\%93_Soling&oldid=945362535)
using a Plackett-Luce extension of the Bradley-Terry model.

@@

```{r}
(soling_diff <- soling_after - soling)
```

>>>>

---
title: "soling"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Soling is a type of boat raced in the 2000 Summer Olympics.  Here I
analyse race data taken from
[wikipedia](https://en.wikipedia.org/w/index.php?title=Sailing_at_the_2000_Summer_Olympics_\%E2\%80\%93_Soling&oldid=945362535)
using a Plackett-Luce extension of the Bradley-Terry model.

@@

```{r solingdiff}
(soling_diff <- soling_after - soling)
```

### 2023-10-30 inst/table_tennis_serve.Rmd
consistentification

---
title: "The importance of serve in table tennis"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{table_tennis_serve}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# The importance of serve in table tennis.

(In this vignette, object `H` is created.  It is identical to
object `table_tennis` in the package; to see it type
`data(table_tennis)` at the command line, and to see the help
page, type `help(table_tennis)`).

Suppose we have three players, "A","B", and "C" who play table tennis.
We seek to quantify the effect of the serve.  The scoreline is as
follows:

>>>>

---
title: "The importance of serve in table tennis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
In this vignette, object `H` is created.  It is identical to object
`table_tennis` in the package; to see it type `data(table_tennis)` at
the command line, and to see the help page, type `help(table_tennis)`.
Suppose we have three players, "A","B", and "C" who play table tennis.
We seek to quantify the effect of the serve.  The scoreline is as
follows:

### 2023-10-30 inst/tennis.Rmd
consistentification

---
title: "Tennis doubles and ghosts"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document defines and analyses a small dataset originally
considered by Hankin (2010).  There, results from tennis games among
four players was considered:

@@

The likelihood for the 9-2 scoreline between $\left\lbrace
p1,p2\right\rbrace$ and $\left\lbrace p3,p4\right\rbrace$ would have a
likelihood function of

@@

* RKS Hankin 2010. "A generalization of the Dirichlet distribution". 
 _Journal of Statistical Software_ 33(11)
* RKS Hankin 2017. "Partial rank data with the hyper2 package:
  likelihood functions for generalized Bradley-Terry models".  _The R
  Journal_ 9(2)

>>>>

---
title: "Tennis doubles and ghosts"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document discusses a small dataset first presented by
@hankin2010.  There, results from tennis games among four players was
considered:

@@

The likelihood for the 9-2 scoreline between $\left\lbrace
p_1,p_2\right\rbrace$ and $\left\lbrace p_3,p_4\right\rbrace$ would
have a likelihood function of

@@

### 2023-10-30 inst/universities.Rmd
name all chunks and cache the slow ones
consistentification

---
title: "universities"
output: bookdown::html_document2
---

@@

## University ranking data

First load the data:

```{r cars}
universities_table <- read.table("universities.txt",header=TRUE)
head(universities_table,12)
```

@@

```{r,label=preandpost}

@@

```{r}
H <- H1 + H2
maxp(H)
```

@@

```{r}
samep.test(H,c("AUT_pre","AUT_post"))
```

>>>>

---
title: "university ranking data"
output: bookdown::html_document2
author: "R. K. S. Hankin"
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I analyse a dataset on NZ university rankings, taken from the
Times Higher Educational Supplement.  First load the data:

```{r loadthetable}
uni_table <- read.table("universities.txt",header=TRUE)
universities_table <- uni_table
head(universities_table,12)
```

@@

```{r,label=preandpost,cache=TRUE}

@@

```{r maxph1h2,cache=TRUE}
H <- H1 + H2
maxp(H)
```

@@

```{r sameptesthon2,cache=TRUE}
samep.test(H,c("AUT_pre","AUT_post"))
```

### 2023-10-30 inst/volleyball.Rmd
name all chunks and cache the slow ones
consistentification

---
title: "volleyball dataset"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```
<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document discusses a dataset first presented by Hankin
(2010), although here only the first 52 observations are used.  A
volleyball \dfn{set} is a Bernoulli trial between two disjoint subsets
of the players.  The two subsets are denoted (after the game) as the
_winners_ and the _losers_: these are denoted by `1` and `0`
respectively.

@@

Each row of `volleyball_table` is a set.  Thus the first line shows a
game between a team comprising `p1`, `p4`, and `p8` against a team
comprising `p5` and `p6`.  Player `p9` did not play; team `p1 p4 p8`
won.

We may use function `volley()` to convert this to a likelihood
function:

```{r calcmaxpvolley}
volleyball <- volley(volleyball_table)
(volleyball_maxp <- maxp(volleyball))
```

>>>>

---
title: "volleyball dataset"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

```{r setup, include=FALSE}
set.seed(0)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
options("digits" = 5)
```

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document discusses a dataset first presented by @hankin2010,
although here only the first 52 observations are used.  A
volleyball _set_ is a Bernoulli trial between two disjoint subsets
of the players.  The two subsets are denoted (after the game) as the
_winners_ and the _losers_: these are denoted by `1` and `0`
respectively.

@@

Each row of `volleyball_table` is a set.  Thus the first line shows a
game between a team comprising `p1`, `p4`, and `p8` against a team
comprising `p5` and `p6`.  Player `p9` did not play; team `p1 p4 p8`
won.  We may use function `volley()` to convert this to a likelihood
function:

```{r calcmaxpvolley,cache=TRUE}
volleyball <- volley(volleyball_table)
(volleyball_maxp <- maxp(volleyball))
```

### 2023-10-30 inst/volvo.Rmd
label all chunks, cache the slow ones
consistentification

---
title: "Race results from the twelfth edition of the
  round-the-world Volvo Ocean Race"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document creates the `volvo` suite of R objects and
presents some preliminary analysis.  It refers to the 12th edition of
the Volvo Ocean Race, 2014-2015 in which seven racing teams competed
over nine legs.  Dataset taken from the wikipedia page (see references
below).

@@

```{r}
volvo <- ordertable2supp(volvo_table,noscore=c("DNF","DNS"))
(volvo_maxp <- maxp(volvo))
```

@@

```{r,equalvolvostrengths}
equalp.test(volvo)
```

>>>>

---
title: "Race results from the twelfth edition of the
  round-the-world Volvo Ocean Race"
author: "Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document creates the `volvo` suite of R objects and
presents some preliminary analysis.  It refers to the 12th edition of
the Volvo Ocean Race, 2014-2015 in which seven racing teams competed
over nine legs.  Dataset taken from the wikipedia page (see references
below).

@@

```{r specifynoscore,cache=TRUE}
volvo <- ordertable2supp(volvo_table,noscore=c("DNF","DNS"))
(volvo_maxp <- maxp(volvo))
```

@@

```{r,label=equalvolvostrengths,cache=TRUE}
equalp.test(volvo)
```

### 2023-10-30 inst/zermelo.Rmd
consistentification

---
title: "Some thoughts on Zermelo's iteration process in the context of the hyper2 package"
output: bookdown::html_document2
---

@@

<p style="text-align: right;">
![](`r system.file("help/figures/hyper2.png", package = "hyper2")`){width=10%}
</p>

This short document sets out a use-case for functions `zermelo()` and
`pairwise()`.  First I create `n=10` competitors with Zipf strengths:

>>>>

---
title: "Some thoughts on Zermelo's iteration process in the context of the hyper2 package"
author: "Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

```{r out.width='20%', out.extra='style="float:right; padding:10px"',echo=FALSE}
knitr::include_graphics(system.file("help/figures/hyper2.png", package = "hyper2"))
```

To cite the `hyper2` package in publications, please use @hankin2017a.
This short document sets out a use-case for functions `zermelo()` and
`pairwise()`.  First I create `n=10` competitors with Zipf strengths:
